\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-05-03 trunk/concepts.tex
updates

\eu uses a \emph{domain model} written in a declarative language,
together with initial conditions and goals (also in that specified
language), to construct a set of temporal relations that must be true
at start time. These models include assertions about the physics of
the vehicle, i.e how it responds to external stimulus and internally
driven goals. By propagating these relations forward using Simple
Temporal Networks \cite{dechter91} and applying goal constraints, \eu
can select a set of conditions that should be true in the future,
where some of these conditions will correspond to actions the agent
must take. The planner can backtrack and try another path during the
search process if a goal cannot be reached while being capable of
discarding unachievable goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a variable and contains an explicit network of bounded
delay constraints between such
variables. Fig. \ref{fig:flex-timelines} shows an example of a
flexible partial plan.

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.3]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time, the most appropriate for actual execution. The fact that
constraints are explicitly represented ensures that through constraint
propagation the executive will respect global limits expressed in the
plan (e.g., don't start a task until a certain condition has been
satisfied but still satisfying some overall deadline). Such
flexibility is critical when dealing with dynamic and uncertain
conditions such as in the ocean, where precise timing of a robotic
action might be indeterminate. Further, the advantage of flexibility
can be contrasted with the consequences of the intrinsic inflexibility
of traditional command sequences. Because they are inflexible,
sequences must necessarily be designed considering worst case
scenarios.

\eu uses a \emph{state variable} representation
\cite{mus94,jonsson00,mcgann09} to describe the evolution of state
over time. The instantiated history of such state variable evolution
over a temporal horizon we call \emph{timelines} and which represent a
single thread in the execution of a concurrent system. At any given
time each thread can execute a single procedure.  Thus each timeline
consists of a sequence of procedures which encapsulate and describe
state evolution; we call these instantiated atomic entities
\emph{tokens}.  A token therefore describes a procedure invocation,
the state variables on which it can occur, the parameter values of the
procedure, and the time values defining the interval. We allow
encapsulation of uncertainty within these tokens with a range of start
and end times and parameters, all of which are encoded as variables. A
constraint solver in turn manipulates these variables defined in a \eu
domain model. For example, consider a scientific need to take water
samples 100 meters from an environmental hotspot while an AUV is
performing a Yo-Yo sawtooth pattern in the water-column. Two samples
are needed if the feature's signal is above a threshold; one
otherwise. The token that is capturing the sensory threshold has a
parametric constraints to the token which fires the requisite water
sampler. In addition, the start time of the water sampling procedure
is highly dependant on the variability of sub-sea currents and actual
speed of the vehicle. Therefore a number of values are possible for
the start times and duration of the sampling all of which are valid
combinations for desired outcomes.

@@

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain and over which we may want to specify
  constraints. In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where variables representing time is used.  As noted in
  Section \ref{sec:europa:cp}, a variable can take values from a
  domain. In \eu, domains are defined over a specific data type, the
  primitive data types supported are, \textit{String},
  \textit{Boolean}, \textit{Numeric} and \textit{Object} (see
  below). Further, two types of variable Domains can be defined over a
  data type, \textit{Enumerated} and \textit{Interval} (used only for
  numeric data types, they are specified by a [LowerBound,UpperBound]
  pair, for instance: [1,10], [5.0,100.0]). It is this latter ability
  to represent and reason over numeric intervals that forms the basis
  of representing flexible plans.

\item[\textbf{Constraints}] valid plans in most problem domains have
  to satisfy a number of restrictions, for instance, a Shopping Agent
  may only be able to perform one task at a time, may have limited
  time or an energy or fuel budget to perform a task. In \eu, these
  restrictions are represented by constraints, which can be defined
  over any combination of variables in a domain. \eu provides a
  \emph{Constraint Library} that implements a number of useful
  constraints (temporal, resource, relational), while allowing the
  possibility to add new constraints if required by a particular
  domain.

\item[\textbf{Objects}] The items we wish to describe and refer to in
  a domain are considered objects. As in the case with object-oriented
  analysis and design, one can seek out the nouns in any domain
  description to find likely objects. In the Shopping Agent example,
  we might consider the Agent, Products and Locations all to be
  objects. Objects have state and behavior, for example, a Shopping
  Agent can have, its location, a bag that contains the products it
  has already purchased (the bag can in turn be another object) and
  behavior as in going to a location to look for a product, perform a
  purchase, return home, etc.

  Objects that have similar state and behavior can be described
  generically in terms of object types or classes. \eu allows the
  definition of classes similar to object-oriented programming
  languages, where a class can have attributes (represented by
  variables), and classes can be arranged in a single-parent hierarchy
  through inheritance.  However, one key aspect in \eu that is not
  found in object-oriented programming: to describe state and behavior
  for the purposes of planning we need to build on the formalism of
  first order logic.

\item[\textbf{Tokens}] In first order logic, a predicate defines a
  relation between objects and properties \cite{gen87}.  In \eu, we
  define such relations between variables whose domains are sets of
  objects and sets of properties to describe state and behavior. For
  example, one might use a predicate \texttt{At($a$,$l$)} to indicate
  that agent $a$ is at location $l$, or a predicate
  \texttt{Buy($a$,$p$}) to indicate that agent $a$ is taking action to
  buy product $p$. Note that $a$ is a variable which may have a number
  of possible values in a problem with multiple agents. Similarly, $l$
  and $p$ are variables whose values are the set of possible positions
  and products respectively. \eu can be used to create partial plans,
  where the domains of variables $a$ and $l$ can have more than one
  possible value. Or they can be grounded plans, where single values
  will be specified for each variable as showing in the N-queen's
  problem.

  In general, for an executable plan it is insufficient to state
  predicates that describe required state or behavior without also
  specifying some temporal extent over which each of those predicates
  hold. Predicates that are always true can be thought to hold from
  the beginning to the end of the planning horizon. However, in
  practice, the temporal extent of interest must be defined with
  timepoints \cite{Dean88} to represent its start and end.  In such
  representation, one writes \texttt{At($a$,$s$,$e$,$l$)} to indicate
  that agent $a$ is at location $l$ from time $s$ to time $e$. In
  fact, this pattern of using such predicates to describe both state
  and behavior of objects is prevalent enough that \eu the token
  representation. Token have built in variables to indicate the object
  to which the statement principally applies as well as the timepoints
  over which it holds. A token is an instance of a predicate that
  represents an object's state or behavior and is defined over a
  temporal extent. In \eu, all predicate instances are tokens. Also,
  in the same way that objects are described generically by classes,
  in \eu tokens can be described generically by token types. Every
  token has five built-in variables:

@@

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  tokens.  Goals can be posted explicitly, for example, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  rules. For instance, when the Shopping Agent must get to a location
  where a product is available before it is able to purchase it: in
  this case the original goal of purchasing the product spawns a
  subgoal of being at a specific location without which the purchase
  will not occur as defined in the domain model.  Looking at how a
  simple Shopping Agent goal would be addresed by \eu helps illustrate
  how Token States support the planning process.  Let's assume the
  planner starts with a partial plan that places the Shopping agent at
  home at time $0$ with no products in possession. Let's also assume
  that goals are posted for the Agent to own a drill by time $10$ and
  to be home by time $20$ as shown in Fig. \ref{fig:europapr1}

  \begin{figure}[t]
    \centering \subfloat[\small Timelines representing initial state
    and goals for a Shopping
    Agent.]{\label{fig:europapr1}\includegraphics[scale=0.35]{figs/europa-pr-1.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to be
    home by time $20$ on merging with initial
    state.]{\label{fig:europapr:merge}\includegraphics[scale=0.35]{figs/europa-pr-merge.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to own
    a drill by time $10$ by activating corresponding
    token.]{\label{fig:europapr:activate}\includegraphics[scale=0.35]{figs/europa-pr-activate.pdf}}
    \caption{\small Evolution of a timeline from the Shopping Agent example.}
  \end{figure}

  Note that the tokens \texttt{Own(Drill)} and \texttt{At(Home)} that
  represent the goals are not immediately placed on the
  \texttt{AgentBag} or \texttt{AgentLocation} timelines since \eu must
  first determine whether it can achieve those goal and how. When the
  goals are posted and before any decisions are made, the
  \texttt{Own(Drill)} and \texttt{At(Home)} tokens are created in an
  \texttt{INACTIVE} state. \eu then considers the partial plan and
  evaluates two possible alternatives:

@@

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible. Two
tokens are merge-compatible if they are instances of the same
predicate and there are no empty intersections between corresponding
token variables.  The effect of merging is illustrated in
Fig. \ref{fig:europapr:merge}. The goal to be home by time $20$ is
satisfied by merging with the initial state that said that the Agent
is home at time $0$.  The merge is possible because the tokens are of
the same type \texttt{At}, and intersecting the domains for all of
their variables doesn't yield an empty domain (which represents an
inconsistency), in this case the token variables are the location,
which is \texttt{Home} in both cases, and the temporal variables
(start, end, duration), which can be intersected without causing an
inconsistency. After the token representing the goal is merged onto
the token representing the initial state, the resulting token
represents not only the fact that the Agent must be \texttt{At(Home)}
by time 0 but also \texttt{At(Home)} at time $20$, this is reflected
in the new constraints on the token's start and end variables as
depicted in Fig. \ref{fig:europapr:merge}.

When a token is activated, all domain rules associated with its type
are evaluated, which may lead to new subgoals being generated. In our
example, there could be a domain rule stating that if a Shopping Agent
doesn't already own a product, then it must \texttt{Buy} it before it
can \texttt{Own} it. In such a case, activating the
\texttt{Own(Drill)} token will cause a new \texttt{Buy(Drill)} to be
created in an \texttt{INACTIVE} state, which in turn could spawn its
own subgoals, like being at the location where the product can be
purchased before being able to buy it, and so on, until all domain
rules are satisfied. As Fig. \ref{fig:europapr:activate} shows the
decision to merge the goal to be \texttt{At(Home)} by time $20$ with
the initial state was not appropriate since the Agent will have to get
away from \texttt{Home} to be able to buy the drill which is
inconsistent with the constraints that resulted after the merge. As a
result that decision will have to be retracted at some point and the
alternative to activate or merge with a different token will have to
be considered.

If, after looking at all possibilities, neither activating nor merging
work, a token may be \texttt{REJECTED}. This is possible only for
goals posted by the end user and specified as optional. Subgoals
generated as a result of token activation cannot be rejected. If a
subgoal cannot be satisfied through activation or merging then it will
cause the original activation decision that spawned it to be
retracted.

As can be seen from the above example, during the search for a plan, a
token's state can go from \texttt{INACTIVE} to \texttt{ACTIVE, MERGED}
or \texttt{REJECTED}, as shown in Fig. \ref{fig:europapr:states}.

>>>>

\eu uses a \emph{domain model} \kcomment{(see Fig. \ref{fig:planner})}
written in a declarative language, together with initial conditions
and goals (also in that specified language), to construct a set of
temporal relations that must be true at start time. These models
include assertions about the physics of the vehicle, i.e how it
responds to external stimulus and internally driven goals. By
propagating these relations forward using Simple Temporal Networks
\cite{dechter91} and applying goal constraints, \eu can select a set
of conditions that should be true in the future, where some of these
conditions will correspond to actions the agent must take. The planner
can backtrack and try another path during the search process if a goal
cannot be reached while being capable of discarding unachievable
goals.

Traditionally robot execution has relied on dispatching commands at
precise times. Such linear sequences of precisely timed commands give
no ability to adjust execution on the basis of sensory
information. Although some commands can issue tests on sensor
readings, these tests have the objective of verifying whether expected
execution conditions are occuring. If not, the state of the system is
declared off-nominal and execution of the sequence of commands is
interrupted. More recently, executives have been proposed and
implemented that significantly broaden the way robots can be commanded
\cite{mus98,alami:1998p820}. For example, the Remote Agent executive
interpreted a \textit{temporally flexible plan} which represents each
start time as a \kcomment{flexible timepoint} variable
\kcomment{backed by} an explicit network of bounded delay constraints
between such variables. \kcomment{These timepoints \cite{Dean88}
  represent temporal intervals signifying a change in state and are of
  the form $[lb,ub]$ (where $lb, ub \in N$) are temporal lower and
  upper bounds respectively. Instead of specifying a fixed integer,
  time is represented as an interval}. Fig. \ref{fig:flex-timelines}
shows an example of a flexible partial plan.

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.4]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\end{figure}

Unlike a traditional fixed time-tagged command sequences, such
flexible plans leave room for adaptation at execution time. When the
executive considers when to start a task, it propagates information
through the constraint network, computes a time bound for the
variable, selects an actual execution time within the bound, and
starts the task at that time. Temporally flexible plans therefore,
express a \textit{range of possible outcomes} of the robots
interaction with the environment within which the executive can elect
at run time, the most appropriate for actual execution. The fact that
constraints are explicitly represented ensures that through constraint
propagation the executive will respect global limits expressed in the
plan (e.g., don't start a task until a certain condition has been
satisfied but still satisfying some overall deadline). Such
flexibility is critical when dealing with dynamic and uncertain
conditions \kcomment{for robotic action in real-world environments},
where precise timing of an action might be indeterminate. Further, the
advantage of flexibility can be contrasted with the consequences of
the intrinsic inflexibility of traditional command sequences. Because
they are inflexible, sequences \kcomment{are brittle and therefore}
must necessarily be designed considering worst case scenarios.

\eu uses a \emph{state variable} representation
\cite{mus94,jonsson00,py10} to describe the evolution of state over
time. The instantiated history of such \kcomment{an} evolution over a
temporal horizon we call \kcomment{as a} \emph{timeline} and
\kcomment{one which represents} a single thread in the execution of a
concurrent system. At any given time each thread can execute a single
procedure\kcomment{, with} each timeline consisting of a sequence of
procedures which encapsulate and describe state evolution. We call
these instantiated atomic entities \emph{tokens}.  A token therefore
describes a procedure invocation, the state variables on which it can
occur, the parameter values of the procedure, and the
\kcomment{timepoints} defining the interval \kcomment{(see
  Fig. \ref{fig:flex-timelines})}. We allow encapsulation of
uncertainty within these tokens with a range of start and end times
and parameters, all of which are encoded as variables. A constraint
solver in turn manipulates these variables defined in an \eu domain
model. For example, consider a scientific need to take water samples
100 meters from an environmental hotspot while an AUV is performing a
Yo-Yo sawtooth pattern\footnote{\kcomment{Variability in the water
    column is along the vertical dimension. Since the Dorado platform
    can only move forward, the Yo-Yo pattern is the most efficient
    mechanism for studying water column properties.}}  in the
water-column. Two samples are needed if the feature's signal is above
a threshold; one otherwise. The token that is capturing the sensory
threshold has a parametric constraints to the token which
\kcomment{triggers} the requisite water sampler. In addition, the
start time of the water sampling procedure is highly dependant on the
variability of sub-sea currents and actual speed of the
vehicle. Therefore a number of values are possible for the start times
and duration of the sampling all of which are valid combinations for
desired outcomes.

@@

\item[\textbf{Variables}] Values that need to be represented to
  describe the problem domain \kcomment{over which constraints are
    specified}.  In the Shopping Agent example, the times at which the
  agent needs to leave or be back, or executes a purchase, are all
  instances where \emph{variables} representing time is used.  As
  noted in Section \ref{sec:europa:cp}, a variable can take values
  from a domain. In \eu, domains are defined over a specific data
  type; the primitive data types supported are \textit{String},
  \textit{Boolean}, \textit{Numeric} and \textit{Object} (see
  below). Further, two types of variable Domains can be defined over a
  data type \textit{Enumerated} and \textit{Interval} (used only for
  numeric data types, they are specified by a
  $[LowerBound,UpperBound]$ pair, for instance: $[1,10]$,
  $[5.0,100.0]$). It is this latter ability to represent and reason
  over numeric intervals that forms the basis of representing flexible
  plans.

\item[\textbf{Constraints}] Valid plans in most problem domains have
  to satisfy a number of restrictions for instance, a Shopping Agent
  may only be able to perform one task at a time, may have limited
  time or an energy or fuel budget to perform a task. These
  restrictions are represented by \emph{constraints}, which can be
  defined over any combination of variables in a domain. \eu provides
  a \emph{Constraint Library} that implements a number of useful
  constraints (temporal, resource, relational) while allowing the
  possibility to \kcomment{augment} with new constraints if required.

\item[\textbf{Objects}] The items we wish to describe and refer to in
  a domain are considered objects.  In the Shopping Agent example, we
  might consider the Agent, Products and Locations all to be
  \emph{objects}. Objects have state and behavior.  For example, a
  Shopping Agent can have its location, a bag that contains the
  products it has already purchased (the bag can in turn be another
  object) and behavior, as in going to a location to look for a
  product, perform a purchase, return home, etc. all as objects.
  Those that have similar state and behavior can be described
  generically in terms of object types or classes. \eu allows the
  definition of classes similar to object-oriented programming
  languages, where a class can have attributes (represented by
  variables) and classes can be arranged in a single-parent hierarchy
  through inheritance.  \kcomment{What is different however, is that
    to describe state and behavior for planning one needs} to build on
  the formalism of first order logic.

\item[\textbf{Tokens}] In first order logic, a predicate defines a
  relation between objects and properties \cite{gen87}.  In \eu, we
  define such relations between variables whose domains are sets of
  objects and sets of properties to describe state and behavior. For
  example, one might use a predicate \texttt{At($a$,$l$)} to indicate
  that agent $a$ is at location $l$, or a predicate
  \texttt{Buy($a$,$p$}) to indicate that agent $a$ is acting to buy
  product $p$. Note that $a$ is a variable which may have a number of
  possible values in a problem with multiple agents. Similarly, $l$
  and $p$ are variables whose values are the set of possible positions
  and products respectively. \eu can be used to create partial plans,
  where the domains of variables $a$ and $l$ can have more than one
  possible value. Or they can be grounded \kcomment{representations},
  where single values will be specified for each variable as shown in
  the $N$-Queens problem.

  In general, for an executable plan it is insufficient to
  \kcomment{define} predicates that describe required state or
  behavior without also specifying some temporal extent over which
  each of those predicates hold. Predicates that are always true can
  be thought to hold from the beginning to the end of the planning
  horizon. However, in practice, the temporal extent of interest must
  be defined with timepoints to represent its start and end.  In such
  \kcomment{a} representation, one writes \texttt{At($a$,$s$,$e$,$l$)}
  to indicate that agent $a$ is at location $l$ from time $s$ to time
  $e$. In fact, this pattern of \kcomment{usage} to describe both
  state and behavior of objects is prevalent enough that \eu
  \kcomment{encapsulates into} the token
  representation. \kcomment{Tokens} have built in variables to
  indicate the object to which \kcomment{it primarily} applies as well
  as the timepoints over which it holds. A token, therefore, is an
  instance of a predicate that represents an object's state or
  behavior and is defined over a temporal extent. In \eu, all
  predicate instances are tokens. \kcomment{In} the same way that
  objects are described generically by classes, in \eu tokens can be
  described generically by token types. Every token has five built-in
  variables:

@@

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  tokens.  Goals can be posted explicitly, for example, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a consequence of
  domain rules. For instance, the Shopping Agent must get to a
  location where a product is available before it is able to purchase
  it. In this case the original goal of purchasing the product spawns
  a \emph{subgoal} of being at a specific location without which the
  purchase will not occur as defined in the domain model.

  Looking at how a simple Shopping Agent goal would be addresed by \eu
  helps illustrate how Token States support the planning process.
  Let's assume the planner starts with a partial plan that places the
  Shopping agent at home at time $0$ with no products in
  possession. Let's also assume that goals are posted for the Agent to
  own a drill by time $10$ and to be home by time $20$ as shown in
  Fig. \ref{fig:europapr1}

  \begin{figure}[t]
    \centering 
    \subfloat[\small Timelines representing initial state
    and goals for a Shopping Agent.]{\label{fig:europapr1}\includegraphics[scale=0.4]{figs/europa-pr-1.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to be
    home by time $20$ on merging with initial state.]{\label{fig:europapr:merge}\includegraphics[scale=0.4]{figs/europa-pr-merge.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to own
    a drill by time $10$ by activating corresponding
    token.]{\label{fig:europapr:activate}\includegraphics[scale=0.4]{figs/europa-pr-activate.pdf}} 
    \caption{\small Evolution of a timeline from the Shopping Agent example.}
  \end{figure}

  Note that the tokens \texttt{Own(Drill)} and \texttt{At(Home)} that
  represent the goals are not immediately placed on the
  \texttt{AgentBag} or \texttt{AgentLocation} timelines, since \eu
  must first determine whether it can achieve those \kcomment{goals}
  and how. When the goals are posted and before any decisions are
  made, the \texttt{Own(Drill)} and \texttt{At(Home)} tokens are
  created in an \texttt{INACTIVE} state. \eu then considers the
  partial plan and evaluates two possible alternatives:

@@

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target \texttt{ACTIVE} token that is
compatible. Two tokens are merge-compatible if they are instances of
the same predicate and there are no empty intersections between
corresponding token variables.  The effect of merging is illustrated
in Fig. \ref{fig:europapr:merge}. The goal to be home by time $20$ is
satisfied by merging with the initial state that \kcomment{indicated}
that the Agent is home at time $0$.  The merge is possible because the
tokens are of the same \texttt{At} type and intersecting the domains
for all of their variables does not yield an empty domain (representing
an inconsistency). In this case the token variables are the location,
which is \texttt{Home} in both cases and the temporal variables start,
end and duration, which can be intersected without causing an
inconsistency. After the token representing the goal is merged onto
the token representing the initial state, the resulting token
represents not only that the Agent must be \texttt{At(Home)} by time
$0$ but also \texttt{At(Home)} at time $20$, This is reflected in the
new constraints on the token's start and end variables as depicted in
Fig. \ref{fig:europapr:merge}.

When a token is activated, all domain rules associated with its type
are evaluated, which may lead to new subgoals being generated. In our
example, there could be a domain rule stating that if a Shopping Agent
doesn't already own a product, then it must \texttt{Buy} it before it
can \texttt{Own} it. In such a case, activating the
\texttt{Own(Drill)} token will cause a new \texttt{Buy(Drill)} to be
created in an \texttt{INACTIVE} state, which in turn could spawn its
own subgoals, like being at the location where the product can be
purchased before being able to buy it, and so on, until all domain
rules are satisfied. As Fig. \ref{fig:europapr:activate} shows the
decision to merge the goal to be \texttt{At(Home)} by time $20$ with
the initial state was not appropriate since the Agent will have to get
away from \texttt{Home} to be able to buy the drill which is
inconsistent with the constraints that resulted after the merge. As a
result that decision will have to be retracted and the alternative to
activate or merge with a different token will need to be considered.

If after looking at all possibilities, neither activating nor merging
work, a token may be \texttt{REJECTED}. This is possible only for
goals posted by the end user and specified as optional. Subgoals
generated as a result of token activation cannot be rejected. If a
subgoal cannot be satisfied through activation or merging then it will
cause the original activation decision that spawned it to be
retracted.

As can be seen from the above example, during the search for a plan, a
token's state can go from \texttt{INACTIVE} to \texttt{ACTIVE, MERGED}
or \texttt{REJECTED} as shown in Fig. \ref{fig:europapr:states}.
